A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-14 tex/codatadp.tex
more

  * to evaluate procedure, we generated problems from Isabelle theories using
    Sledgehammer

  * chose a selection of theory files:
    * pretty much every theory file from Isabelle distribution and AFP that
      defines a codatatype
    * also included two private theories by Andreas Lochbihler and Peter Gammie
    * (codatatypes being a recent addition to Isabelle \cite{itp},
       they are not yet widely used)
    * and included a number of theories on datatypes
    * aimed at a good mixture between linear structures (lists, streams) and
      tree structures
    * the theory files were chosen before running the experiments

  * how many problems
  * how many problems have some codatatypes, how many have datatypes (including
    XXX that have both)

@@

{\footnotesize
\begin{verbatim}
So we can show the success rates, in a table like this:

                 Codatatype suite     Mixed suite        Datatype suite
                (2345 problems)    (XXX problems)       (1240 problems)

   w/o (co)data        800           1122                699
   w codata            815 (2)       1139 (3)
   w data                            1153 (6)            712 (4)
   w both                            1170 (3+6)
\end{verbatim}}

>>>>

To evaluate the decision procedure, we generated problems from existing
Isabelle formalizations using Sledgehammer \cite{paulson-blanchette-2010}.
Codatatypes being a recent addition to Isabelle
\cite{blanchette-et-al-2014-impl}, benchmarks are somewhat hard to come by. We
included the theory files from the Isabelle distribution and the \emph{Archive
of Formal Proofs} (AFP) \cite{klein-et-al-afp} that define codatatypes falling
within the supported fragment. We also included two unpublished theories by
Andreas Lochbihler and Peter Gammie. We complemented the benchark suite with
some theories about various list and tree data structures, to also test the
datatypes. The theory files were selected before running the experiments.

@@

\newcommand\landg{L\,\&\,G}
\newcommand\HD[1]{\hbox to2.5em{\hfill#1\hfill}}

\begin{table}[t!]
\begin{center}\begin{tabular}{l@{\kern1.5em}c@{\kern.5em}c@{\kern.5em}c@{\kern1.5em}c}
  & \HD{Distro} & \HD{AFP} & \HD{\landg} & All
\\[1pt] %%% TYPESETTING HACK
\midrule
%$\enatT$
\\[-9pt] %%% TYPESETTING HACK
No decision procedure
  & 800 & 922 & 450 & 2172 \\
Codatatypes without bisimilarity
  & 810 & 936 & 470 & 2216 \\
Codatatypes with bisimilarity
  & 811 & 939 & 472 & 2222 \\
Datatypes without cyclicity
  & 815 & 947 & 469 & 2231 \\
Datatypes with cyclicity
  & 818 & 953 & 470 & 2241 \\
Full decision procedure
  & 840 & 970 & 490 & 2300 \\[\jot]
Total number of goals
  & 1211\phantom{0} & 1121\phantom{0} & 899 & 3001
\end{tabular}\end{center}
\caption{Number of solved goals for three benchmark suites}
\end{table}

Do we lose any problems by enabling it?
